### 2020-03-04 doc/icfp20/harmony.tex
minor tweaks to harmony

The primary Example~9.1 in Piston's text
(Figure~\ref{fig:harmonization} left) shows a
sample four part SATB harmonization in C~major. The melody is the
series of highest notes, i.e., the soprano (S) line, while the other three
parts alto (A), tenor (T), and bass (B) are the melodies below
it. Roman numerals denote the names of the triads at each vertical
slice. Note that even though there are four voices there are only
three distinct pitches sounding at each vertical slice (where tones an
octave apart are considered to have the same relative pitch). As is
typical the root of the triad is always doubled. Also note that all
triads are in root position (meaning the root is in the bass voice),
as Piston restricts to this case initially in the chapter (he does
relax it later). The goal of a harmonization is both to create an
aesthetically pleasing chord progression as well as fluid melodies in
each voice (called voice leading) and independence of each music
line (in other words good counterpoint). In this chapter
Piston gives advice on how to achieve these goals.

@@

In Haskell this would be easy---there is an \texttt{Eq} typeclass and a
method for automatically generating equalities. In the Agda Standard
Library~\citep{agda-stdlib} list membership is a relation with
constructive evidence of membership, and requires far more effort we
would like given that we currently don't need to do
proofs. Alternatives include defining decidable equality for triads
(which requires writing out 49 cases) or mapping to natural numbers
and using already implemented decidable or boolean equality for
those. For now we simply write the inverse function manually,
excluding the rare cases.

@@

We next restrict to progressions ending in a half cadence (V) to match
Piston's example (\texttt{filter halfCadence}). This results in 25
candidate progressions. For each pitch and triad pair we construct a
\texttt{harmonizingChord} in root position with the root doubled. The
function \texttt{voiceChord} spaces the notes nicely. All of this work
ins done in the function \texttt{voicedHarmonizations}.

@@

To determine the ``best'' of these harmonizations, we now make use of
\texttt{motionCheck} from the counterpoint work. We compare all six
pairs of melodies, count the number of motion errors present, and pick
the harmonization which minimizes these flaws. The functon to
determine all errors is straightfoward.

>>>>

The primary Example~9.1 in Piston's text
(Figure~\ref{fig:harmonization} left) shows a
sample four part SATB harmonization in C~major. The melody is the
series of highest notes, i.e., the soprano (S) line, while the other three
parts alto (A), tenor (T), and bass (B) are the melodies below
it. Roman numerals denote the names of the triads at each vertical
slice. Note that even though the example consists of four voices there are only
three distinct pitches sounding at each vertical slice (where tones an
octave apart are considered to have the same relative pitch). As is
typical the root of the triad is always doubled. Also note that all
triads are in root position (i.e., the root is in the bass voice),
as Piston restricts to this case initially in the chapter (he does
relax it later). The goal of a harmonization is both to create an
aesthetically pleasing chord progression as well as fluid melodies in
each voice (called voice leading) and independence of each music
line (in other words good counterpoint). In this chapter
Piston gives advice on how to achieve these goals.

@@

In Haskell this would be easy---there is an \texttt{Eq} type class and a
method for automatically generating equalities. In the Agda Standard
Library~\citep{agda-stdlib} list membership is a relation with
constructive evidence of membership, and requires far more effort we
would like given that we currently don't need to do
proofs. Alternatives include defining decidable equality for triads
(which requires writing out 49 cases) or mapping to natural numbers
and using already implemented decidable or boolean equality for
those. For now we simply write the inverse function manually,
excluding the rare cases.

@@

We next restrict to progressions ending in a half cadence (V) to match
Piston's example (\texttt{filter halfCadence}). This results in 25
candidate progressions. For each pitch and triad pair we construct a
\texttt{harmonizingChord} in root position with the root doubled. The
function \texttt{voiceChord} spaces the notes nicely. All of this work
is done in the function \texttt{voicedHarmonizations}.

@@

To determine the ``best'' of these harmonizations, we now make use of
\texttt{motionCheck} from the counterpoint work. We compare all six
pairs of melodies, count the number of motion errors present, and pick
the harmonization which minimizes these flaws. The function to
determine all errors is straightforward.
